\section{Cyclotomic Polynomials}

\begin{definition}
The cyclotomic polynomial $\phi(x)$ for non-trivial $p$-th roots of unity (other than $x = 1$) is given by
\[ \phi(x) = {(x^p - 1)\over x - 1} = x^{p-1} + x^{p - 2} + \hdots x + 1. \]
\end{definition}

\begin{theorem}
$\phi(x)$ is an irreducible polynomial.

\end{theorem}
\begin{proof} As it stands,  the Eisenstein criterion for irreducibility doesn't apply. 
However a change of variable $y = x - 1$ and the binomial expansion changes the form into 
\[ {(x^p - 1)\over x - 1}  = {((y+1)^p - 1)\over y}  = y^{p-1} + p y^{p-2} + {p(p-1)\over 1\cdot2} y^{p-3}  + \hdots + p.\]
The binomial coefficients which appear on the right hand side are all integers divisible by the prime $p$, for $p$ occurs in each numerator as a factor, and can never be cancelled out by the smaller integers in the denominator.
The polynomial in $y$ therefore satisfies the Eisenstein criterion and is therefore irreducible -- which means that the original equation for $\phi(x)$ is irreducible.  \qed
\end{proof}

\begin{theorem}[Irreducibility of $x^p - a = 0$]
A polynomial $f(x) = x^p - a = 0$  of prime degree $p$ over a field $K$ is either irreducible over $K$ or has a root in $K$.
\end{theorem}

\begin{proof}[Proof 1] {\it based on the Rational Root Theorem}
Let $\alpha$  be a root of $f(x) = x^p - a$ (assume irreducible or not). Let $K = \Q(\alpha)$.
(Assume we're working over $\Q$ or $\Z$ by Gauss's lemma for monic polynomials).
\begin{itemize}
\item[]Since $p$ prime, $[K:\Q] =1 \mbox{ or } p$.
\item[]If =1, $\alpha \in \Q$, so rational root.
\item[]Hence if no rational root, $[K:\Q]=p$, so minimum polynomial has degree $p$, so $f$ (monic) is the minimum polynomial, hence irreducible.
\end{itemize}
\end{proof}

\begin{proof}[Proof 2]
Adjoin to $K$ a primitive $p$-th root of unity $\xi$ and the a root $u$ of $x^p - a$. The resulting extension $K(\xi, u)$ contains the $p$ roots 
$u, \xi u, \xi^2 u, \hdots, \xi^{p-1} u$ of the polynomial $x^p - a$, hence is the root field of this polynomial, which has the factorization
\[ x^p - a = (x-u)(x-\xi u)(x - \xi^2 u) \hdots (x-\xi^{p-1} u). \]
Suppose that $x^p -a$ has a proper factor $g(x)$ over $K$ of positive degree $m < p$. This factor $g(x)$ is then a product of $m$ of the linear factors of $x^p - a$ over $K(\xi,u)$, 
so that the constant term $b$ in $g(x)$ is a product of the roots $\xi^i u$. Therefore $b = \xi^k u^m$, for some integer $k$, and 
\[ b^p = (\xi^k u^m)^p = (\xi^p)^k (u^p)^m = (u^p)^m = a^m. \]
From this we can find in $K$ a $p$-th root of $a$, for $m<p$ is relatively prime to $p$ and there exists integers $s$ and $t$ with $sm + tp = 1$, so that  
\[ b^p = a^m = a^{1 -tp} = a/a^{tp}, \]
and
$a = (b^s a^t)^p$. The assumed reducibility of $x^p -a$ over $K$ therefore yields a root $(b^s a^t)$ of $x^p - a$ in K. \qed
\end{proof}

\begin{theorem}[Abel's Version os the Irreducibility of $x^p - F$]
If the equations 
\begin{eqnarray}
f_1 x^{p-1} + f_2 x^{p-2} + \hdots + f_p &=& 0 \cr
x^p - F &=& 0
\end{eqnarray}
where $p$ is a prime number, are simultaneously satisfied, either $f_1, f_2, f_3, \hdots, f_p$ all vanish, or else one of the roots of $x^p - F = 0$ can be expressed rationally in terms of $f_1, f_2, \hdots, f_p$ and $F$.
\end{theorem}

\begin{proof}
For, suppose the coefficients $f_1, f_2, \hdots, f_p$ don't vanish. 
The the two equations will have a greatest common divisor (GCD) of the form
\[ g(x) = x^\rho + g_1 x^{\rho - 1} + g_2 x^{\rho - 2} + \hdots + g_\rho = 0, \]the coefficients of which are 
rational functions of $F, f_1, f_2, \hdots, f_p$.
Now if $x_1$ be any one of the roots common to both equations, the other roots will be of the form 
\[ \xi^\alpha x_1, \xi^\beta x_1, \hdots. \mbox{ where } \xi^\rho -1 = 0; \]
from which 
\[ g_\rho = x_1^\rho \xi^{\alpha+\beta}\hdots = \xi^\tau x_1^\rho. \] The rest of the 
proof  with small changes follows Proof 2 above -- since $p$ is a prime number, we can find two numbers $m$ and $n$, of which one is negative, such that 
$mp + n\rho = 1$. Also, 
\[ g_\rho^n = \xi^{n\tau}x_1^{n\rho} = \xi^{n\tau} x_1^{(1-m\rho)}, \] and so 
\[ \xi^{n\tau}x_1 = g_\rho^n F^m, \] and therefore $\xi^{n\tau}x_1$, which is a root of $x^p - F = 0$, is expressed rationally in terms of $F, f_1, f_2, \hdots, f_p$.   
\end{proof}