\section{Invariant Subgroups}

Invariant Subgroups as also known as Normal Subgroups. \\

Consider a subgroup $H$ of $G$. For $g, g' \in G$ let $x = gH$ and $y = g'H$ be the corresponding cosets of $g$ and $g'$. For $H$ to be an Invariant Subgroup of $G$ we require
that $xy$ to be in the same coset as $gg'H$ or \[(gH) (g'H) \in gg'H.\] 

To obtain the basic relation defining Invariant Subgroups, premultiply both sides by $g^{-1}$:
\begin{eqnarray*}
g^{-1} (gH) (g'H) &\in& g^{-1}gg'H, \cr
H (g' H) &\in& g'H. \cr
\end{eqnarray*} 
In the right most $H$ choose the neutral element as $e$ 
\begin{eqnarray*}
Hg' e &\in& g'H \cr
Hg'   &\in& g' H \cr
\end{eqnarray*}  If $Hg'\in g'H$ for any $g'\in G$ we also must have that (changing $g'$ to $g$) that $Hg \in gH$. Since $g^{-1} \in G$ we also have 
\[ Hg^{-1} \in g^{-1} H.\] and therefore
\begin{eqnarray*}
 (gH)g^{-1}g &\in& gg^{-1} Hg,\cr
 gH &\in& Hg,
 \end{eqnarray*}
 Therefore we have $gH \in Hg$ and $Hg \in gH$ and so \[ gH = Hg,\] leading up to the following definition of an Invariant or Normal Subgroup: 
\begin{definition}
$H$ is an invariant subgroup of $G$ if it is a subgroup and if the left and right cosets of $H$ in $G$ coincide; i.e., if 
\[ gH = Hg \quad \forall g \in G.\]
\end{definition}

Invariant subgroups are often called normal subgroups or self-conjugate subgroups. \\

If the left coset $gH$ is also a right coset, then it contains  $g (= eg)$ and so must be the right coset $Hg$. Hence if every left coset is a right coset (or vice versa) $H$ is invariant.\\

Let us show that this condition is sufficient.
\begin{theorem}
If $H$ is invariant in $G$ then $(gH) (g'H) = g g' H$ for all $g,g' \in G$. 
\end{theorem} 

\begin{proof}
\begin{eqnarray*}
 gH g'H &=& g (Hg') H = g(g'H)H \mbox{ since } Hg' = g'H  \mbox{ as H is invariant,}\cr
 &=& gg'H^2 = gg'H \mbox{ since } H^2 = H \mbox{ as } H \mbox{ is a subgroup.}
\end{eqnarray*}
\end{proof}
Note we have already shown that it is sufficient to prove that \[ (gH)(g'H) \subseteq gg'H,\] in order to show that $H$ is an Invariant Subgroup. 

\begin{theorem}
A subgroup $H$ is an invariant subgroup of $G$ is and only if $g^{-1}Hg = H$ for all $g\in G$. 
\end{theorem}

\section{Conjugacy}
\begin{definition}
The elements $x$ and $g^{-1} x g$, where $x$ and $g$ are elements of a group $G$, are called conjugate elements in $G$ : in other words $x$ and $y$ are conjugate if and only if there exists
and element $g \in G$ such that $y = g^{-1} x g$.
\end{definition}

\begin{theorem}
The relation of conjugacy between elements of any group $G$ is an equivalence relations. 
\begin{itemize}
\item[]{\it ~~~~~Reflexive.} Since $x = e^{-1} x$, $x$ is conjugate to itself.
\item[]{\it ~~~~~Symmetric.} If $ y = g^{-1} x g$ we have $x = g y g^{-1} = (g^{-1})^{-1} y g^{-1}$. 
\item[]{\it ~~~~~Transitive.}  If $  y = g^{-1} x g$ and  $z = h^{-1} y h$ we have \[ z = h^{-1}g^{-1} x g h = (gh)^{-1} x (gh).\] 
\end{itemize}
\end{theorem}
It follows that the relation divides the elements of $G$ into mutually exclusive classes; these are called {\it conjugacy classes in G}.

\begin{theorem}
If $H$ is a subgroup of $G$ then $g^{-1} H g$ is a subgroup for any element $g\in G$. 
\end{theorem}
\begin{proof}
For \[(g^{-1} H g)^2 = (g^{-1} H g g^{-1} H g) = (g^{-1} H H g) = (g^{-1} H g) \mbox{   since }  H^2 = H\] as $H$ is a subgroup. Also
\[  (g^{-1} H g)^{-1} = g^{-1} H^{-1} g = g^{-1} H g  \mbox{   since } H^{-1} = H.\]  
\end{proof}

We say that $H$ and $g^{-1} H g$ are {\it conjugate subgroups} in $G$. 

\begin{theorem}
The relation of conjugacy between subgroups of $G$ is an equivalence relation.
\end{theorem}

\begin{theorem}
Normal subgroups of $G$ are composed of all elements of $g$ which are common to the set of conjugate subgroups of $G$. 
\end{theorem}
\begin{proof}
Suppose the set of conjugate subgroups of $G$ are
\[ H_1, H_2, \hdots H_i.\]
The elements which are common to these subgroups form a group, since with any two of them contained in $H_i$ their product is also contained in $H_i$. 
Also, the transformations of $H_1$ by elements in $G$ are the same conjugate subgroups
\[ H_1, g_2^{-1}H_1 g_2, \hdots, g_i^{-1} H_1 g_i, \] the group of common elements, which is part of any one of them as it is of $H_1$, is identical with its 
transforms and hence a normal subgroup of $G$. 
\end{proof}
A group of all elements which are common to conjugate subgroups, is called their {\bf greatest common subgroup} and a greatest common subgroup 
is normal in the group. 

\begin{theorem}
$H$ is an invariant subgroup of $G$ if the conjugacy class of subgroups of $G$ which contains $H$ consists of just $H$ itself: i.e., if the only subgroup
conjugate to $H$ is $H$. 
\end{theorem}