\section{Substitutions and Almost Symmetric Functions}

Substition refers to the set of permutations of $n$ elements which form the Symmetric Group.

\begin{theorem}
All the substitutions of $x_1, x_2, \hdots, x_n$ which leave unaltered a rational function
$\phi(x_1, x_2,\hdots, x_n$ form a group $G$.
\end{theorem}

Let $\phi_s$ denote the function obtained by applying to $\phi$ the substitusion $s$. If $a$ and $b$
are two substitutions which leave $\phi$ unaltered, then $\phi_a = \phi, \phi_b = \phi$. Then

\[ (\phi_a)_b = (\phi)_b = \phi_b = \phi, \mbox{ or } \phi_{ab} = \phi \]
Hence the product of $ab$ is one of the substitutions which leave $\phi$ unaltered and the set
$G$ has the group property. \\

The group $G$ is called the {\it group of the function $\phi$}, while $\phi$ is said to {\it belong to the
group $G$}\\

\begin{theorem}
Every substitution can be expressed as a product of transpositions in various ways.
\end{theorem}

\begin{theorem}
Of the various decompositions of a given substitution $s$ into a product of transpositions, all contain
an even number of transpositions (an even substitution), or all constain an odd number of transpositions
(an odd substitution).
\end{theorem}

\begin{corollary}
The totality of even substitutions on $n$ letters forms a group, called the Alternating Group on n letters.
\end{corollary}


Consider $n$ elements $x=\{ x_1, x_2, x_3, \hdots, x_n \}$. Consider the rational function

\[
\begin{array}{r}
\phi(x) = \prod(x_i - x_j) = (x_1-x_2) (x_1-x_3) (x_1-x_4) \quad \hdots\hdots\quad (x_1 - x_n) \\
(x_2-x_3) (x_2-x_4) \quad \hdots\hdots\quad (x_2 - x_n) \\
(x_3-x_4) \quad \hdots\hdots\quad (x_3 - x_n) \\
\quad\quad \hdots\hdots\hdots \quad\quad\quad~~\\
(x_{n-1} - x_n) \\
\end{array}
\]
consisting of the product of all the difference of the elements.\\

$\phi(x)$ is related to the Vandemonde determinant

$$\phi(x) = V(x) =
\begin{vmatrix}
1 & x_1 & x_1^2 & x_1^3 & \hdots & x_1^{n-1} \\
1 & x_2 & x_2^2 & x_2^3 &\hdots &x_2^{n-1} \\
1 & x_3 & x_3^2 & x_3^3 &\hdots & x_2^{n-1} \\
\vdots \\
1 & x_n & x_n^2 & x_n^3 &\hdots & x_n^{n-1} \\
\end{vmatrix}
$$

The square of $V(x)$ is a
well- known symmetric function, the discriminant $\triangle$. Therefore $V(x)$ has two values equal
numerically with opposites signs, viz. $\sqrt{\triangle}$ and $-\sqrt{\triangle}$. $V(x)$ is an example
of an alternating function. Any transposition alters the sign of $V(x)$. For example, the transposition
$(x_1x_2)$ affects only the terms in the first and second lines of the product, and replaces them by
\[
\begin{array}{r}
(x_2-x_1) (x_2 - x_3) (x_2 - x_4) \quad \hdots\hdots\quad (x_2 - x_n) \\
(x_1-x_3) (x_1-x_4) \quad \hdots\hdots\quad (x_1 - x_n). \\
\end{array}
\]
Hence, if $s$ is the product of an even number of transpositions, it leaves $v(x)$ unaltered; if $s$
is the product of an odd number of transpositions, it changes $V(x)$ into $-V(x)$. \\

{\bf Conjugate values of Multiple-valued Function and Conjugate Groups}\\

\begin{theorem}[Lagrange] The order of a subgroup is a divisor of the order of the group.
\end{theorem}

\begin{proof}
Consider a group $G$ of $N$ and a subgroup $H$ composed of the substitutions
\be h_1 = I, h2, h3, \hdots, h_P. \label{HL1} \ee
If G constains no further substitutions, then $N = P$ and the theorem is true. Now let $G$ contain a
substitution $g_2$ not in $H$. Then G also must contain
\be g_2, h_2 g_2, h_3 g_2, \hdots, h_P g_2. \label{HL2} \ee This set are all distinct and all different from the substitutions
(\ref{HL1}), since $h_\alpha g_2 = h_\beta$ requires that $g_2 = h_\alpha^{-1} h_\beta = $ a substition of $H$
contrary to the assumption that $g_2$ is distinct from the elements of $H$. Hence the substitutions combined now
give $2P$ distinct substitutions of $G$. If there are no other substitutions in $G$, then $N=2P$ and the theorem
is true. If there are any other distrinct substitutions $g_3$ in $G$ not in (\ref{HL1}) or (\ref{HL2}), then $G$ contains
\be g_3, h_2 g_3, h_3 g_3, \hdots, h_P g_3. \label{HL3} \ee Moreover, these substitutions are all different from (\ref{HL1})
and (\ref{HL2}), since $h_\alpha g_3 = h_\beta g_2$ requires that $g_3 = h_\alpha^{-1}h_\beta g_2$ and shall belong
to (ref{HL2}), contrary to assumption. We now have $3P$ distinct substitutions of $G$. Either $N=3P$ or else $G$
contains a substitution $g_4$ not in the sets (\ref{HL1}), (\ref{HL3}), (\ref{HL3}). In the latter case $G$ contains the products
\be g_4, h_2 g_4, h_2 g_4, \hdots, h_P g_4, \label{HL4} \ee all of which are distrinct and all different from the sets
(\ref{HL1}), (\ref{HL3}), (\ref{HL3}), so that we have $4P$ distinct substitutions. Proceeding in this way, we finally
reach a last set of $P$ substitutions
\be g_nu, h_2 g_nu, h_3 g_nu, \hdots, h_P g_\nu, \label{HLP} \ee since the order of$ H$ is finite. Hence $N=\nu P$.
\end{proof}


\begin{definition}
The number $\nu$ is called the {\it index} of the subgroup $H$ under $G$ and is denoted by $\nu = [G \mathbin{:} H]$.
\end{definition}

\begin{corollary}
The order of any group $H$ of substitutionis on $n$ letters is a divisor of $n!$. $H$ is a subgroup of the symmetric group $G_n$
on $n$ letters.
\end{corollary}

\begin{theorem}
The period of any substitution contained in a group $G$ of order $N$ is a divisor of $N$.
\end{theorem}

\begin{corollary}[Applied to the Symmetric Group]
The order of any substitution group on a set of $n$ elements is an exact divisor of $N=n!$, the quotient
being the number of distinct value of the corresponding multiple-valued function.
\end{corollary}

Let us arrange the $N$ substitutions of$G$ in a rectangular array using the substitutions of a subgroup $H$
in the first row:

\be
\begin{array}{lllll}
h_1=I & h_2 & h_3 & \hdots & h_P \\
g_2 & h_2 g_2 & h_3 g_2 & \hdots & h_P g_2 \\
g_3 & h_2 g_3 & h_3 g_3 & \hdots & h_P g_3 \\
\vdots & \hdots & \hdots & \hdots & \vdots\\
g_\nu & h_2 g_\nu & h_3 g_\nu & \hdots & h_P g_\nu \\
\end{array}\label{Lagrange}
\ee

The $g_1 = I, g_2, g_3, \hdots, g_\nu$ multiply the elements of $H$ on the right hand side forming right cosets.
Similarly, a rectangular array for the substitiions could be formed by using left costs.

\begin{theorem}
If $\psi$ is a rational funciton of $x_1, x_2, \hdots, x_n$ belonging to a subgroup $H$ of index $\nu$, then $\psi$
is $\nu$-valued under the full set of substitutions of $G$.
\end{theorem}

\begin{proof}
Apply to $\psi$ all the $N$ substitutions of $G$ arranged in a rectangular array as in (\ref{Lagrange}). All the substitutions
belonging to the same row give the same answer since
\[ \psi_{h_i g_\alpha} = (\psi_{h_i} )_{g_\alpha} = (\psi)_{g_\alpha} = \psi_{g_\alpha}. \]
Hence there result at most $\nu$ values. But, if
\[ \psi_{g_\alpha} = \psi_{g_\beta} \]
then $\psi_{g_\alpha g_beta^{-1}} = \psi,$ so that $g_\alpha g_\beta^{-1} h_i \in H$, leaving $\psi$ unaltered. Hence $g_\alpha = h_i g_\alpha$, contrary to the assumption made
in forming the rectangular array.
\end{proof}

Regarding the orders of a group $H$ and a larger group $G$ containing $H$ as a subgroup; that is to say, the order $r$ of $H$ is an exact divisor of the order $G$ which contains $H$ as a subgroup, the quotient $m$ being the number of distinct values of a multiple-valued function unaltered by the substitutions of $G$ which are obtained by the substutions of $H$.

\begin{definition}
The $\nu$ distinct functions $\psi, \psi_{g_2}, \psi_{g_3}, \hdots, \psi_{g_\nu}$ are called {\bf conjugate values} of $\psi$ under $G$.
\end{definition}

\begin{theorem}
The $\rho$ distinct values which a rational function $\psi(x_1, \hdots, x_n)$ takes when operated on by all $n!$ substitutions are the roots of
an equation of degree $\rho$ whose coefficients are ration functions of the elementary symmetric functions.

\end{theorem}

{\bf Formation of Functions of a Given Group}\\

We define the Galois Function having $n!$ values for all the substiutionts of the symmetric group
\[ \psi_1 = \alpha_1 x_ 1 + \alpha_2 x_ 2 + \alpha_3 x_ 3 + \hdots \alpha_n x_ n,\] in which $\alpha_1, \alpha_2, \hdots, \alpha_n$ are $n$ distinct arbitrary constants. This function is
called the Galois Function. If we use all the $P$ substitutions of $H$ to form the functions $\psi_1, \psi_2, \hdots, \psi_P$ by the substitutions of $H$ anysymmetric funciton of these values will be unaltered by the substitutions of $H$. In particular $\phi_1 = (y-\psi_1)(y-\psi_2)\hdots(y-\psi_P)$ will be unaltered by the substutions of $H$, and altered to a different value by the substitutions not in $H$. The function $\phi_1$ is not therefore unaltered by the substitutions of any wider group that contains $H$ as a subgroup. Noting the forms of the expression of the sums of powers of the roots in terms of the coefficients, we see that instead of the coefficients of the powers of $y$ in $\phi_1$ we may take sums of powers of $\psi_1, \psi_2, \hdots, \psi_P$ up to the $P^{th}$ and deduce that one at least of such $P$ sums of powers will provide a function unlatered by the substitutions of $H$ and altered by any other substitution of the symmetric group.

\begin{theorem}
Every integral symmetric function of the distinct values of any integral multiple-valued function of $n$ elements is a symmetric function of the elements themselves.
\end{theorem}

\begin{proof}
Let $F(\phi_1, \phi_2, \hdots, \phi_\rho)$ be any rational integral symmetric function of the $\rho$-values. Any substitution $S$ whatever affecting the elements applied to these $\rho$-values
either leaves any function unchanged or replaces it by one of the others. No two of the resulting values can be equal, for if $\phi_i S$ were equal to $\phi_j S$ it would follow, by applying the substitution $S^{-1}$,  that $\phi_i = \phi_j$ which is contrary to the assumptions. Consequently the $\rho$ values of $\phi$ are reproduced by $S$ in some order or other. The symmetric function $F$ therefore remains unchanged by any substitutiont, and is consequently a symmetric element of the elements themselves. 
\end{proof}

\begin{corollary}
The $\rho$ distinct values of an integral multiple-valued function are roots of an equation whose coefficients are integral symmetric functions of the elements themselves.
\end{corollary}

\begin{theorem}[Lagrange Theorem on functions belonging to the same group]
Two rational functions belonging to the same group are rationally expressed each in terms of the other.
\label{FunctionsOfSameGroup}\end{theorem}

\begin{proof}
Let $\phi_1$ and $\psi_1$ be two functions belonging to the same group
\[ H = \{ I, S_2, S_3, \hdots, S_r \}, \] of order $r$ and degree $n$, each of these functions having $\rho$ distinct values, where $r*\rho = N$. Any substitution not contained in $H$
will convert $\phi_1$ into another of its values, say $\phi_k$ and at the same time convert $\psi_1$ into $\psi_k$. By operating all possible substitutions, $\rho$ pairs of values
$\phi_1, \psi_1 : \phi_2, \psi_2 : \hdots \phi_\rho, \psi_\rho$ are obtained. Now, in the first place, the rational function

\be \sum \phi^i \psi^j = \phi_1^i \psi_1^j + \phi_2^i \psi_2^j + \hdots + \phi_k^i \psi_k^j + \hdots + \phi_\rho^i \psi_\rho^j \label{psiphi}\ee
is clearly a symmetric function of the elements, for it appears by the same reasoning as before that any substitution acting on this function will merely permute the terms leavng the sum unaltered. Therefore $\sum phi^i \psi^j$ is a symmetric function of the elements. \\

If we now take $j = 1$ and assign to $i$ all the values $0,1,2,\hdots,\rho-1$ in succession, we obtain the following $\rho$ equations linear in $\psi_1, \psi_2, \hdots, \psi_\rho$:
\be
\begin{array}{r c r c r c r r}
\psi_1 & + & \psi_2 & + & \hdots & + & \psi_\rho = & T_0 \\
\phi_1 \psi_1 & + & \phi_2\psi_2& + & \hdots & + &\phi_\rho \psi_\rho = & T_1 \\
\phi_1^2 \psi_1 & + & \phi_2^2\psi_2& + & \hdots & + &\phi_\rho^2 \psi_\rho = & T_2 \\
\hdots & \hdots & \hdots & \hdots & \hdots & \hdots &\hdots = & \hdots \\
\phi_1^{\rho-1} \psi_1 & + & \phi_2^{\rho-1}\psi_2& + & \hdots & + &\phi_\rho^{\rho-1} \psi_\rho = & T_{\rho-1} \\
\end{array}\label{linearSys}
\ee where $T_0, T_1, \hdots, T_{\rho-1}$ are all symmetric functions of $x_1, x_2, \hdots, x_n$.
\end{proof}
We can write the above as a linear system
\be
\begin{bmatrix}
1 & 1 & 1 & 1 & \hdots & 1 \\
\phi_1 & \phi_2 & \phi_3 & \phi_4 & \hdots & \phi_\rho \\
\phi_1^2 & \phi_2^2 & \phi_3^2 & \phi_4^2 & \hdots & \phi_\rho^2 \\
\vdots & \vdots & \vdots& \vdots & \vdots & \vdots \\
\phi_1^{\rho-1} & \phi_2^{\rho-1} & \phi_3^{\rho-1} & \phi_4^{\rho-1} & \hdots & \phi_\rho^{\rho-1} \\
\end{bmatrix}
\begin{bmatrix}
\psi_1 \\
\psi_2 \\
\vdots\\
\vdots\\
\psi_\rho
\end{bmatrix}
=
\begin{bmatrix}
T_0 \\
T_1 \\
\vdots\\
\vdots\\
T_{\rho-1}
\end{bmatrix}
\label{LagrangeSystem}\ee

Linear systems such as Eq.(\ref{LagrangeSystem}) can be solved using Cramer's rule. Alternatively, in order to understand better the relatonship between $\phi_i$ and $\psi_i$ let us consider
a method based on transforming non-homogenous linear systems into sets of corresponding homogeneous linear systems in each of the unknown parameters. To this effect consider 
the linear system \cite{Burnside}.
\be
\begin{array}{r r r r r r r r}
x               & + &             y   & + &                z     & = & A_0\\
\alpha x    & + &     \beta y  & + & \gamma z      & = & A_1 \\
\alpha^2 x & + & \beta^2 y  & + & \gamma^2 z  & = & A_2 \\
\end{array}
\label{ExampleLinearSystem}
\ee
By adding a fourth equation to the above we can isolate one of the variables, say $y$, and determine the functional dependence of $y$ on $\beta$ and the symmetric functions of the coefficients.
Adding $y = y$ to the above system and converting it into a system of homogenous equations by introducing a 4th unknown $w$ we obtain the system 

\be
\begin{array}{r r r r r r r r}
0               & + &             y   & + &                0     & = & y   \\
x               & + &             y   & + &                z     & = & A_0\\
\alpha x    & + &     \beta y  & + & \gamma z      & = & A_1 \\
\alpha^2 x & + & \beta^2 y  & + & \gamma^2 z  & = & A_2 \\
\end{array}
\ee

The above system can be converted into a Homogeneous System by moving the final column to the left-hand side of the equal sign. 

\be
\begin{array}{r r r r r r r r r r}
0               & + &             y   & + &                0     & - & y    & = & 0\\
x               & + &             y   & + &                z     & - & A_0& = & 0\\
\alpha x    & + &     \beta y  & + & \gamma z      & - & A_1 & = & 0\\
\alpha^2 x & + & \beta^2 y  & + & \gamma^2 z  & - & A_2 & = & 0\\
\end{array}
\ee

Writing the above system using a matrix of coefficients and introducing the 4th undetermined variable we find
\be
\begin{bmatrix}
0            & 1           & 0 & -y  \\
1            & 1           & 1 & -A_0  \\
\alpha    & \beta     & \gamma & -A_1  \\
\alpha^2 & \beta^2 & \gamma^2 & -A_2  \\
\end{bmatrix}
\begin{bmatrix}
x \\
y \\
z \\
w \\
\end{bmatrix}
=
\begin{bmatrix}
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\label{BandP}\ee

In order for this homogenous system to be consistent with Eq.(\ref{ExampleLinearSystem}) the determinant of the matrix of coefficients must vanish. 
\begin{equation}
\begin{vmatrix}
0            & 1           & 0               & -y    \\
1            & 1           & 1               & -A_0\\
\alpha    & \beta     & \gamma    & -A_1\\
\alpha^2 & \beta^2 & \gamma^2 & -A_2\\
\end{vmatrix}\end{equation}
Since the determinant vanishes we can factor out the $-1$ from the last column and write the consistency condition as 

\begin{equation}
\begin{vmatrix}
0            & 1           & 0               &  y    \\
1            & 1           & 1               &  A_0\\
\alpha    & \beta     & \gamma    &  A_1\\
\alpha^2 & \beta^2 & \gamma^2 & A_2 \\
\end{vmatrix} = 0.
\end{equation}

Multiplying the above determinant by the non-singular determinant 
\begin{equation}
\begin{vmatrix}
1            & 1           & 1               &  0  \\
\alpha    & \beta     & \gamma    &  0   \\
\alpha^2 & \beta^2 & \gamma^2 & 0   \\
0            & 0           & 0               &  1   \\
\end{vmatrix} 
\end{equation}we find

\begin{equation}
\begin{vmatrix}
0            & 1           & 0               &  y     \\
1            & 1           & 1               &  A_0 \\
\alpha    & \beta     & \gamma    &  A_1 \\
\alpha^2 & \beta^2 & \gamma^2 & A_2  \\
\end{vmatrix} 
\begin{vmatrix}
1   & \alpha    & \alpha^2   &  0    \\
1   & \beta      & \beta^2    &  0    \\
1   & \gamma & \gamma^2 & 0    \\
0    & 0           & 0               &  1    \\
\end{vmatrix}
=
\begin{vmatrix}
1            & \beta     & \beta^2      &  y    \\
s_0        & s_1       & s_2             &  A_0\\
s_1        & s_2       & s_3             &  A_1\\
s_2        &s_3        & s_4             & A_2 \\
\end{vmatrix} = 0,
\label{BandPFinal}\end{equation}
where \[s_i = \alpha^i + \beta^i + \gamma^i\] and the $s_i$ are symmetric functions in $\alpha, \beta, \mbox{  and } \gamma$. \\ 

Since the $s_i$'s are sums of powers of $\alpha, \beta, \gamma$ and are symmetric functions, they can be expressed rationally in terms of the elementary symmetric functions. By expanding the final determinant in Eq. (\ref{BandPFinal}) along the top row and setting the expansion equal to zero to satisfy the consistency requirement we see that we have $y$ expressed as a quadratic function of $\beta$ along with the sums of powers of the three quantities $\alpha, \beta, \gamma$. 

Returning to the original problem  Eq. (\ref{linearSys}), we can similarly express the solution for $\psi_j$ in terms of $\phi_j$ in the form of a determinant: 

\be
\begin{vmatrix}
1            & \phi_j     & \phi_j^2      &  \hdots & \phi_j^{\rho-1} & \psi_j  \\
s_0        & s_1        & s_2      &  \hdots & s_{\rho-1}              & T_0     \\
s_1        & s_2        & s_3      &  \hdots & s_\rho                     & T_1     \\
\hdots    & \hdots    & \hdots        &  \hdots & \hdots              & \hdots  \\
s_{\rho-1}            & s_\rho     & s_{\rho+2}      &  \hdots & s_{2\rho-2} & T_{\rho-1}   \\
\end{vmatrix} = 0,
\ee where $T_0, T_1, \hdots, T_{\rho-1}$ are all symmetric functions of $x_1, x_2, \hdots, x_n$ and $s_k = \phi_1^k + \phi_2^k + \phi_3^k + \hdots + \phi_\rho^k$.\\

The solution of these equations takes the form 
\be \triangle(\phi_1, \phi_2, \phi_3, \hdots, \phi_\rho) \psi_j = A_0 \phi_j^{\rho-1} + A_1 \phi_j^{\rho-2} + \hdots + A_{\rho-1},\ee
where $\triangle$ is the product of the square of the differences of all the $\phi$'s (the square of the Vandermonde determinant) and $A_0, A_1, \hdots, A_{\rho-1}$ are all symmetric in $x_1, x_2, \hdots, x_n$. \\

Therefore $\psi_j$ can be rationally expressed as a rational function of $\phi_j$ and the elementary symmetric functions of $x_1, x_2, \hdots, x_n$ (the coefficients of the polynomial that has $x_1, x_2, \hdots, x_n$ as roots). \\

{\bf Lagrange Interpolation Formula: } \\
Given a set of $k+1$ nodes $\{x_0, x_1, \hdots, x_k\}$, which must all be distinct, $x_j \ne x_m$ for indices $j\ne m$, the {\it Lagrange basis} for polynomials of degree $\le k$ for those nodes is the set of polynomials $\{ l_0(x), l_1(x), \hdots, l_k(x)\}$ each of degree $k$ which take values $l_j(x_m) = 0$ if $m\ne j$ and $l_j(x_j) = 1$. Using the Kronecker delta this can be written
$l_k(x_m) = \delta_{jm}$. Each basis polynomial can be explicitly described by the product: 

\begin{eqnarray*}
l_j(x) &=& {(x-x_0)\over (x_j-x_0)}{(x-x_1)\over (x_j-x_1)} \hdots {(x-x_{j-1})\over (x_j-x_{j-1})}{(x-x_{j+1})\over (x_j-x_{j+1})}\hdots{(x-x_k)\over (x_j-x_k)}\\
         &=& \prod_{\substack{0\le m\le k \\ m \ne j }}\,\, {x-x_m\over x_j - x_m}
\end{eqnarray*}

Notice that the numerator $\displaystyle \prod_{m\ne j} (x-x_m)$ has $k$ roots at the nodes $\{x_m\}_{m\ne k}$ while the denominator 
$\displaystyle \prod_{m \ne j} (x_j -x_m)$ scales the resulting polynomial so that $l_j(x_j) = 1$. \\

The Lagrange interpolating polynomial for those nodes through the corresponding values $\{ y_0, y_1, \hdots, y_k\}$ is the linear combination
\[ L(x) = \sum_{j=0}^k \, y_j l_j(x),\]
Each basis polynomial has degree $k$, so the sum $L(x)$ has degree $\le k$, and it interpolates the data because 
\[ L(x_m) = \sum_{j=0}^k y_j l_j(x_m) = \sum_{j=0}^k y_j \delta_{mj} = y_m,\]

\begin{theorem}[Extension of Lagrange's Theorem to Non Symmetric Functions] If two rational functions of any set of variables are such that one remains unchanged by all the substitutions of the group to
which the other belongs, the first is expressible by means of the second in the form of an integral polynomial whose coefficients are rational symmetric functions of the variables. \\

{\bf Another statement of Lagrange's Theorem: } \\

If a rational function $\phi(x_1,x_2, \hdots, x_n)$ remains unaltered by all the substitutions which leae another rational function $\psi(x_1, x_2, \hdots, x_n)$ unaltered, then $\phi$ is
a rational function of $\psi$ and $c_1, c_2, \hdots, c_n$. Note that the group, $G$ to which $\phi$ belongs contains the group, $H$ to which $\psi$ belongs, i.e. $H\subseteq G$. $G$ is the
wider group. Also $phi$ is therefore the {\it more symmetric} function -- a wider set of substitutions keep it unaltered.
\end{theorem}

\begin{proof}
\be
\begin{array}{llll | r | r}
I & h_2 & \hdots & h_P & \psi = \psi_1 & \phi = \phi_1 \\
g_2 & h_2 g_2 & \hdots & h_P g_2 & \psi_{g_2} = \psi_2 & \phi_{g_2} = \phi_2 \\
\hdots & \hdots & \hdots & \hdots & \hfill \hdots \hfill & \hfill \hdots \hfill\\
g_\nu & h_2 g_\nu & \hdots & h_P g_\nu & \psi_{g_\nu} = \psi_\nu & \phi_{g_\nu} = \phi_\nu \\
\end{array}\
\ee
The $\psi_1, \psi_2, \hdots, \psi_\nu$ are all distinct; but$\phi_1, \phi_2, ...\phi_nu$ need not be distinct since $\phi$ belongs to a group $G$ which may be bigger than $H$.
Under any substitution $s$ on $x_1, x_2, \hdots, x_n$, the functions $\psi_1, \psi_2, \hdots, \psi_\nu$ are merely permuted. Moreover, if $s$ replaces
$\psi_i$ by $\psi_j$, it replaces $\phi_i$ by $\phi_j$. Set (using the Lagrange Interpolation formula)

\begin{eqnarray*}
g(t) &=& (t-\psi_1)(t-\psi_2)\hdots, (t-\psi_\nu), \cr
\lambda(t) &=& g(t) \left( {\phi_1\over t- \psi_1} +{\phi_2\over t- \psi_2} + \hdots + {\phi_\nu\over t- \psi_\nu} \right)
\end{eqnarray*}so that $\lambda(t)$ is an integral function of degree $\nu-1$ in $t$. Since $\lambda(t)$ remains unaltered under every substitution of $s$, its coefficients
are rational symmetric functions of $x_1, x_2, \hdots, x_n$ and hence are rational functions of the elementary symmetric functions. \\

Now we're assuming here that $x_1, x_2, \hdots, x_n$ denote indeterminant quantities since $\psi_1, \psi_2, \hdots, \psi_nu$ are algebraically distinct, so that $g'(\psi)$ is not idendically
zero. In the case where special values are assigned to $x_1, x_2, \hdots, x_n$ such that two or more of the functions $\psi_1, \psi_2, \hdots, \psi_nu$ become numerically equaly, then
$g'(\psi) = 0$, and $\phi$ is not a rational function of $\psi, c_1, ..., c_n$.
Assuming that we're working with indetermiant $x_1, x_2, \hdots, x_n$, and
taking $\psi_1 = \psi$ for $t$, we get
\[ \lambda(\psi_1) = \left[ (\psi_1 - \psi_2)(\psi_1-\psi_3)\hdots (psi_1-\psi_nu) \right]\phi_1\] and in general
\[ \phi = {\lambda(\psi) \over g'(\psi)}.\]
\end{proof}

\begin{theorem}
Every two valued function of $n$ variables is of the form $S_1 \pm S_2\sqrt{\triangle}$, where $S_1$ and $S_2$ are integral symmetric functions, and $\triangle$ is the discriminant.
\end{theorem}

\begin{proof}
A two-valued function must belong the the group of order $\half N$. The only group of this order is the alternating group to which the function $\sqrt{\triangle}$ belongs. The theorem therefore 
follows as an immediate consequence of the fundamental theorem Thm (\ref{FunctionsOfSameGroup}). \\

{\bf An Second Proof}\\
Let two values of the function be denoted by $\phi_1$ and $\phi_2$ and let $G_1$ and $G_2$ be the corresponding groups, each of order $\half N$. In the first place, these two groups must
be identical; for it any substitution $S$ of $G_1$ were to change $\phi_2$ to its second value $\phi_1$, then $S^{-1}$ would change $\phi_1$ into $\phi_2$; but this is impossible, since $S^{-1}$ and $S$ each belong to $G_1$. Every substitution therefore of $G_1$ must belong to $G_2$ and vice-versa. \\

To show now that these groups coincided with the alternating group, consider the function $\psi = \phi_1 - \phi_2$. Any substitution which belongs to the common group leaves this unaltered; any other will change $phi_1$ to $\phi_2$ and $\phi_2$ to $\phi_1$, and will therefore change the sign of $\psi$; some transposition, $(x_\alpha x_\beta)$ for example, will have this affect, for no group can include all the transpositions without coinciding with the symmetric group. It is easily inferred that $\psi_1 - \psi_2$ is divisible by $x_\alpha - x_\beta$, and hence by the product of all the differences, since $\psi^2$ is symmetric. \\

The quotient of $\psi$ by $\sqrt{\triangle}$ is symmetric. To prove this, let $(\sqrt{triangle})^m$ be the highest power of $\sqrt{\triangle}$ which occurs in $\psi$. The quotient of $\psi$ by
$(\sqrt{\triangle}^m$ is symmetric, since, if not, it would be an alternating function, and would again contain $\sqrt{\triangle}$ as a factor, which is contrary to assumption. It follows immediately that $m$ is an odd number and that the quotient of $\psi$ by $\sqrt{\triangle}$ is symmetric. Writing therefore $\psi_1 - \psi_2 = A\sqrt{\triangle}$ and $\phi_1 + \phi_2 = B$, where $A$ and $B$ are both symmetric, we have 
\[ \phi_1 = S_1 + S_2\sqrt{\triangle} \mbox{  and  } \phi_2 = S_1 - S_2\sqrt{\triangle},\] where $S_1$ and $S_2$ are both symmetric functions of the variables $x_1, x_2, \hdots, x_n$. The groups $G_1$ and $G_2$ obviously coincide with the group of $\sqrt{\triangle}$ -- the Alternatng Group. 
\end{proof}

\begin{theorem}
The alternating functions are the only unsymmetrical functions of $n$ variables of which a power can be symmetric.
\end{theorem}

\begin{proof}
It is sufficient to prove this theorem for prime powers; for if there exists a function $F(x_1, x_2, \hdots, x_n)$ such that $F^{pq}$ is symmetric and $p$ is a prime, then there is 
also a function $\phi = F^q$ such that $\phi^p$ is symmetric. \\

Therefore let \[ \phi^p = S, \mbox{    a symmetric function.}\] Since the group of $\phi$ which is unsymmetric, cannot contain all the transpositions, let $\sigma = (x_\alpha x_beta)$ be a 
transposition which converts $\phi$ into $\phi_j$; we have \[ \phi_j^p = \phi^p = S.\] and therefore $\phi_j = \omega\phi$, where $\omega$ is a $p$th root of unity. Hence
\[ \sigma \phi = \phi_j = \omega \phi,\] and operating again with $\sigma$ 
\[ \sigma^2 \phi = \omega\sigma\phi = \omega^2 \phi,\] but $\sigma^2 = I$ and therefore $\omega^2 = 1$, and consequently $p=2$. \\

Since $\phi^s$ is symmetric, $\phi$ is an alternating function. \qed
\end{proof}

\begin{theorem}
For any number, $n$, of independent elements there is no multiple-valied function of which a power is two-valiued when $n>4$; and when $n=3$, or $n=4$, if there is any such power it is
third power.
\end{theorem}

Concentrating on prime numbers, and supposing that $\phi$ is a multiple-valued function whose $p$ power is two-valued we have
\be \phi^p = S_1 + S_2\sqrt{\triangle}. \label{two-valued}\ee
The group of $\phi$ cannot contain all the circular substitutions of the third order, for if it did, then this group would coincide with the alternating group, and $\phi$ would be two-valued. Let
$\sigma = (x_\alpha, x_\beta, x_\gamma)$ be such a substitutions not contained in the gorup of $\phi$ and suppose $\sigma\phi = \phi_j$. From Eq. (\ref{two-valued}),
since $S_1 + S_2\sqrt{\triangle}$ is unaltered by $\sigma$, we have
\[ \psi^p = \phi_j^p,\] hence $\phi = \omega \phi$, where $\omega$ is a $p^{th}$ root of unity. Operating again twise in succession with $\sigma$ we obtain

\begin{eqnarray*}
\sigma \phi &=& \omega \phi,\cr
\sigma^2 \phi &=& \omega\sigma\phi = \omega^2 \phi,\cr
\sigma^3 \phi &=& \omega^2 \sigma \phi = \omega^3 \phi,\cr
\end{eqnarray*}whence, since $\sigma^3 = 1$ (three cycle), we have $\omega^3 = 1$, and therefore $p=3$.

Again, when the number of elements is greater than 4, there are circular substitutions of the fifth order, and those cannot all be contained in the group of $\phi$ (or else it will be two-valued and belong to the Alternating Group). Let $\tau$ be one of those not contained in this group and $\tau \phi = \phi_j$. We have, as before, from Eq. (\ref{two-valued}), by applying this substitution
(which does not affect the right-hand side),
\[ \psi^p = \phi_j^p = S_1 + S_2\sqrt{\triangle}.\] Hence proceeding as before, we have $\tau\phi = \omega\phi$ and oeprating again on this and the succeeding equations with $\tau$ we
easily find $\tau^5 \phi = \omega^5 \phi$; whence $\omega^5 = 1$, since $\tau^5 = 1$, and it is proved that $p=5$. Now this results being incompatible with the previous value of $p$
obtained, we infer that when the number of elements is greater than 4, it is impossile to find any multiple-valiued function $\phi$, a prime power of which will be two-valued.\\

There are actually, when $n$ is not greater than 4, multiple-valiued functions, a third power of which is two-vaued.
