\section{Vector Spaces in $\R_n$}

\begin{definition}
Vectors are directed segments. The vector in $\R_n$ whose components are $a_1, a_2,...,a_n$ is denoted by $$\a = \{a_1, a_2, ..., a_n\}\hbox{~(braces)}.$$
The zero vector in $\R_n$ is denoted as $$\o = \{0,0,...,0\},$$ with $n$ components each 0. 
\end{definition}

\begin{definition}We call $p$ vectors $\a_1, \a_2, ..., \a_p$ {\bf linearly dependent} if there are $p$ numbers $\lambda_1, \lambda_2,...,\lambda_p$ not all $0$, such that 
$$\lambda_1 \a_1 + \lambda_2 \a_2 +...+ \lambda_p \a_p = \o\hbox{~ (the zero vector)}.$$
\end{definition}

If no such $p$ numbers exist, then the $p$ vectors $\a_1, \a_2, ...,\a_p$ are called {\bf linearly independent}.

\begin{theorem} 
If a subset of the $p$ vectors $\a_1, \a_2,...,\a_p$ is linearly dependent, then the set of $p$ vectors is itself linearly dependent.\label{Th1_1}
\end{theorem}

Suppose the first $r$ vectors $\a_1, \a_2, ..., \a_r, (r < p) $are linearly dependent. Then there are $r$ numbers $\lambda_1, \lambda_2, ..., \lambda_r$ not all zero such that 
$$\lambda_1 \a_1 + \lambda_2 \a_2 + ... + \lambda_r \a_r = \o.$$ Setting $\lambda_{r+1}, \lambda_{r+2}, ..., \lambda_p$ to zero we have
$$\lambda_1 \a_1 + \lambda_2 \a_2 + ... + \lambda_r \a_r + \lambda_{r+1}\a_{r+1} +... + \lambda_p \a_p = \o.$$ where the numbers $\lambda_1, \lambda_2, ..., \lambda_p$ are not all zero. Therefore the 
$p$ vectors $\a_1, \a_2, ..., \a_p$ are linearly dependent. 

\begin{theorem}
If the $p$ vectors $\a_1, \a_2, ..., \a_p$ are linearly independent, then so is every subset of these $p$ vectors. \label{Th1_2}
\end{theorem}

For otherwise, by Theorem \ref{Th1_1}, the $p$ vectors would be linearly dependent.

\begin{theorem}
If the $p$ vectors $\a_1, \a_2, ..., \a_p$ are linearly dependent and if $p>1$, then at least one of these vectors
is a linear combination of the others.\label{Th1_3}
\end{theorem}

For, there are numbers $\lambda_i$, which do not all vanish such that $$\lambda_1 \a_1 + \lambda_2 \a_2 +...+ \lambda_p \a_p = \o.$$ Suppose $\lambda_p \ne 0$. The we can solve for $\a_p$,
$$\a_p = -{\lambda_1\over \lambda_p}\a_1 -{\lambda_2\over \lambda_p}\a_2 - ... - {\lambda_{p-1}\over \lambda_p} \a_{p-1},$$ i.e. $\a_p$ is a linear combination of the vectors $\a_1, \a_2, ..., \a_{p-1}$.  

\begin{theorem}
If one of the vectors $\a_1, \a_2, ..., \a_p$ is a linear combination of the others, then vectors $\a_1,\a_2,...,\a_p$ are linearly dependent.\label{Th1_4}
\end{theorem}

Suppose $$\a_1 = \lambda_2 \a_2 + \lambda_3 \a_3 +... + \lambda_p \a_p.$$ Then we have that 
$$\a_1 - \lambda_2 \a_2 - \lambda_3 \a_3 -... - \lambda_p \a_p = \o,$$ and since the coefficient of $\a_1$ is not zero, the vectors $\a_1, \a_2, ... , \a_p$ are linearly dependent. 

\begin{theorem}
If the vectors $\a_1, \a_2, ..., \a_p$ are linearly independent, and if the vectors $\a_1, \a_2,...\a_p,\b$ are linearly dependent, then $\b$ is a linear
combination of $\a_1, \a_2, ..., \a_p$.\label{Th1_5}
\end{theorem}

We have a relation of the form $$\lambda_1 \a_1 + \lambda_2 \a_2 +... + \lambda_p \a_p + \lambda_{p+1} \b =\o,$$ where not all the $\lambda_i$ vanish. Now
$\lambda_{p+1}$ cannot be 0 because the last term would drop out and then all the other $\lambda_i$ would vanish because of the linear independence of the $\a_i$ vectors. Therefore
$\lambda_{p+1} \ne 0$ and so we have 
$$\b = -{\lambda_1\over \lambda_{p+1}}\a_1 -{\lambda_2\over \lambda_{p+1}}\a_2 - ... - {\lambda_p\over \lambda_{p+1}} \a_p.$$


\begin{theorem}
If the vectors $\a_1, \a_2, ..., \a_p$ are linearly independent, and if $\b$ is a not a linear combination of $\a_1, \a_2, ..., \a_p$, then the vectors
$\a_1, \a_2,...\a_p,\b$ are linearly independent.\label{Th1_6}
\end{theorem}

Theorem \ref{Th1_6} follows from Theorem \ref{Th1_5}.

Let $L$ be a {\elevenit non-empty} set of vectors of $\R_n$ with the properties:
\begin{itemize}
\item[1.] If $\a$ is a vector of the set $L$, then $\lambda \a$ belongs to $L$ for every real $\lambda$.
\item[2.] If $\a$ and $\b$ are two (not necessarily distinct) vectors of the set $L$, then the vector $\a +\b$ also belongs to $L$.
\end{itemize}
Such a set of vectors, satisfying 1. and 2., is  called a {\bf vector space} in $\R_n$.

\begin{theorem} {\bf Steinitz replacement theorem}. Let $\a_1, \a_2, ..., \a_p$ be any $p$ vectors, and let $L$ be the vector space spanned by them. Let $\b_1, \b_2, ..., \b_q$ be any $q$ linearly independent vectors of 
$L$. Then we may replace a certain set of $q$ vectors from among the vectors $\a_1, \a_2, ..., \a_p$ by the vectors $\b_1, \b_2, ...,\b_q$ so that the remaining vectors of $\a_1, \a_2, ..., \a_p$ together with the vectors $\b_1, \b_2, ..., \b_q$ span the entire vector space $L$.\label{Th2_1}
\end{theorem}

\begin{proof}
Use mathematical induction.
\end{proof}

\begin{theorem}
Any set of more than $p$ linear combinations of $p$ given vectors is always linearly dependent.\label{Th2_2}
\end{theorem}

\begin{theorem}
The maximal number of linearly independent vectors in $\R_n$ is $n$\label{Th2_3}
\end{theorem}

\begin{definition}{\bf Dimension and Basis}\\
Let $L$ be any vector space in $\R_n$. The maximal number of linearly independent vectors is $\le n$. Let $p$ be this maximal number of independent vectors in $L$. We call $p$ the {\bf dimension} of $L$. Thus,
$$0 \le p \le n.$$

We call any system of $p$ linearly independent vectors of $L$, such as $\a_1, \a_2, ..., \a_p$ a {\bf basis} of $L$.
\end{definition}

\begin{theorem}
Every vector of $L$ can be represented in exactly one way as a linear combination of the basis vectors $\a_1, \a_2, ..., \a_p$.\label{Th2_4}
\end{theorem}

\begin{theorem}
The dimension of the vector space spanned by the vectors $\a_1, \a_2, ...,\a_p$ is equal to the maximal number of linearly independent vectors among $\a_1, \a_2, ..., \a_p$.\label{Th2_5}
\end{theorem}

\begin{theorem}
Any $k ~ (k\le p)$ linearly independent vectors $$\b_1, \b_2, ..., \b_k$$
 of $L$ can be extended to form a basis of $L$ by suitably adjoining to them $p-k$ other vectors $\a_{k+1}, \a_{k+2}, ..., \a_p$.\label{Th2_6}
\end{theorem}

Let $L'$ and $L''$ be any two vector spaces in $\R_n$. By the {\bf intersection} $D$ of $L'$ and $L''$ we mean the set of all those vectors belonging to both $L'$ and $L''$.\\

If we form the totality $S$ of all vectors of the form $\a' + \a''$, where $\a'$ belongs to $L'$ and $\a''$ belongs to $L''$. $S$ is called the {\bf sum} of $L'$ and $L''$.

\begin{theorem}
If the dimensions of the vector spaces $L', L'', D,$ and $S$ are respectively $p', p'', d$ and $s$, then $$p' + p'' = d + s.$$\label{Th2_7}
\end{theorem}