\section{Vector Spaces}

\begin{definition}
$n$-dimensional Affine Space (Vector Space) $\R_n$. An ordered $n$-tuple of real numbers, i.e., a system of $(a_1, a_2, ..., a_n)$ of $n$ real numbers , called components, in a definite order, is called a point ($n$ a positive integer). The numbers $a_1,a_2, ...,a_n$ are called the coordinates of the point; in particular $a_1$ is called its first, $a_2$ its second, ..., $a_n$ its $n$-th coordinate. Two points $P=(a_1,a_2,...,a_n)$ and $Q = (b_1,b_2,...,b_n)$ are said to 
be the {\elevenit same} or to coincide if and only if $a_1 = b_1, a_2 = b_2, ..., a_n = b_n$. The totality of all $n$-tuples of real numbers is called $n$-dimensional Affine Space and is denoted by $\R_n$.
\end{definition}

\begin{definition}[Vector Notation]
An element, $\x \in \R_n$ is defined by its $n$ components $(x_1, x_2, \hdots, x_n)$, as follows
\[ \x = (x_1, x_2, \hdots, x_n) \]
\end{definition}

\begin{definition}[Linear Dependence/Independence]
In the abstract vector space $V$ the finite set of vectors
$\x_1, \x_2, \hdots, \x_n$ is said to be linear dependent if scalars $\lambda_1, \lambda_2, \hdots, \lambda_n$ exist, not all zero, such that 
\[ \lambda_1 \x_1 + \lambda_2 \x_2 + \hdots + \lambda_n\x_n = \o.\] If no such scalars exist, i.e., if 
\[ \lambda_1 \x_1 + \lambda_2 \x_2 + \hdots + \lambda_n\x_n = \o \implies \lambda_i  = 0 ~\forall ~i = 1,2, \hdots, n\] then the set is linearly independent.  
\label{LinearDependence}
\end{definition}

{\bf Conventions}: We adopt the Range and Summation Conventions:
\begin{definition}
{\bf Range Convention.} {\it When a small Latin suffix (superscript or subscript) occurs unrepeated in a term, it is understood to take all the values
$1,2,...,n$, where n is the number of dimensions of the space.}
\end{definition}

\begin{definition}
{\bf Summation Convention.} {\it When a small Latin suffix is repeated in a term, summation with respect to that suffix is understood, the range of summation
being $1,2,...,n$.}
\end{definition}

{\bf Kronecker delta}:
\[ \delta^s_r = \begin{cases} 1 & \mbox{if } s = r \\0 & \mbox{if } r \ne s\end{cases} \]
Also, the another form of the Kronecker delta is $\delta_{rs}$ which satisfies the same cases as above. \\

{\bf Permutation Symbol}: Consider a set $\sigma$ of $\sigma_1, \sigma_2, ..., \sigma_n$ numbers taken from the set $\{1,2,...,n\}$. Let
\[ \epsilon_{\sigma_1\sigma_2...\sigma_n} =
\begin{cases}
+1 & \mbox{if }\sigma \mbox{ is an even permutation of $\{1, 2, ..., n\}$}\\
-1 & \mbox{if }\sigma \mbox{ is an odd permutation of $\{1, 2, ..., n\}$}\\
0 & \mbox{if }\sigma \mbox{ has any duplicates}
\end{cases} \]

\begin{definition}[Span]
The set of vectors $\x_1, \x_2, \hdots, \x_n$ are said to span or generate, the space $V$ if any vector $\x$ of $V$ is expressible as a linear combination of them, i.e. if 
\[ \x = \lambda_1 \x_1 + \lambda_2 \x_2 + \hdots + \lambda_n \x_n \] for some scalars $\lambda_1, \lambda_2, \hdots, \lambda_n$ not necessarily unique. 
\end{definition}

\begin{definition}[Basis]
The set $\e_1, \e_2, \hdots, \e_n$ is a Basis of $V$: 1) if they are linearly independent; and (2) they span $V$. 
\end{definition}

\begin{theorem}
If $\e_1, \e_2, \hdots, \e_n$ span $V$ then they form a basis if and only if $\x$ in $V$ is expressible uniquely as a linear combination of $\e_1, \e_2, \hdots, \e_n$ .
\end{theorem}

\begin{theorem}
If $\x_1, \x_2, \hdots, \x_m$ span $V$, then there is a subset of these which is a basis of $V$. \label{SpanningSubset}
\end{theorem}
\begin{proof}
If $\x_1, \x_2, \hdots, \x_m$ are linearly independent, they form a basis and the theorem is proved. If they are linearly dependent then some $\x_{k+1}$ is a linear
combination of 
\[  \x_1, x_2, \hdots, \x_k  \mbox{ say } \x_{k+1} = \sum_1^k\, \mu_i \x_i ,  \]
Consider the set $\x_1, \x_2, \hdots, \x_k, \x_{k_2} \hdots, \x_m$. Any vector $\x$ can be expressed as $\displaystyle \x = \sum_1^m \, \lambda_i \x_i$, since the $\x_i$'s also span $V$. 
Hence  
\[ \x = \sum_{\substack{i = 1\\ i\ne k+1j}} \, \lambda_i \x_i  +\lambda_{k+1} \sum_1^k\, \mu_i\x_i  \]
and is a linear combination of $\x_1, \x_2, \hdots, \x_k, \x_{k+2}, \hdots, \x_m$ so that these also span $V$.\\

If this set is now linearly independent the theorem is proved as they form a basis. If not repeat the above argument and obtain a spanning set of $m-2$ vectors, which may or may not
be linearly independent. Repeat until after a most $m-1$ sets we do obtain a basis, which is a subset of the original set. This must occur since after $m-1$ steps we are left with a single vector
which spans $V$, and this must be non-zero unless $V$ is the trivial space $\{ \o \}$ and hence is linearly independent, since $\lambda_1 \x_1 = \o$ with $\x_1 \ne \o \implies \lambda_1 = 0$. 
\end{proof}

\begin{corollary}
If $V$ possesses a finite spanning set then $V$ is finite dimensional. 
\end{corollary}

\begin{theorem}
If $\x_1, \x_2, \hdots, \x_n$ are linearly independent and $V$ is finite dimensional then there exists a basis containing
\[ \x_1, \x_2, \hdots, \x_p.\]
\end{theorem}
\begin{proof}
Let $\e_1, \e_2, \hdots, \e_n$ be a basis of $V$. Consider the set $\x_1, \x_2, \hdots, \x_p, \e_1, \hdots, \e_n$. This set certainly spans $V$. Using the argument of Thm. (\ref{SpanningSubset}) on this set, but notice that 
at each stage we obtain one of the $\e_i$'s as a linear combination of the preceding vectors, since no $\x$ can be a linear combination of the preceding $\x_i$'s as the $\x_i$'s form a linearly independent set. \\

Hence we are left with the $\x_i$'s and the process must stop before or exactly when the $\e_i$'s are exhausted. We this obtain a basis which contains $\x_1, \x_2, \hdots, \x_p$ together with some of the $\e_i$'s. 
\end{proof}

\begin{theorem}[Steinitz Exchange Lemma]
Let $U$ and $W$ be finite subsets of a vector space $V$ and let $U$ be a set of linearly independent vectors and let $W$ span $V$. 
If $\{\x_1, \x_2, \hdots, \x_p\} \in U$ is a linearly independent set, and $\{\y_1, \y_2, \hdots, \y_m\} \in W$ spans $V$, then $p\le m$. \label{TH2_1}
\end{theorem}

\begin{proof}
Consider the set $\x_p, \y_1, \y_2, \hdots, \y_m$, Since the $\y_i$'s span $V$, $\x_p$ is a linear combination of them and so the set is linearly dependent. Hence some $\y_{k+1}$ is a linear combination of 
\[ x_p, \y_1, \hdots, \y_k.\] This means that $\x_p, \y_1, \hdots, \y_k, \y_{k+2}, \hdots, \y_m$ span $V$. \\

Continuing similarly, adding $\x_{p-1}$ to this set. We obtain a linearly dependent set and some $\y_{l_1}$ is expressible ad a linear combination of the others, so that we have a spanning set comprising
$\x_{p-1}, \x_p,$ and all the $\y_i$'s except $\y_{k+1}$ and $\y_{l + 1}$. By continuing this process we obtain a spanning set consisting of one more $\x$ and one less $\y$. We always obtain a $\y_i$ since the $\x_i$'s are 
linearly independent and so no $\x_{i+1}$ is a linear combination of the preceding $\x_i$'s that have been introduced. \\

The process must be possible until all the $\x_i$'s have been exhausted and we must at each stage have some $\y_i$'s left, otherwise the set consisting of the $\x_i$'s along could not be linearly independent. 
Hence we finally obtain a spanning set $\x_1, x_2, \hdots, \x_p$ together with $(m-p)\y_i$'s and thus $m-p \ge 0$. Therefore $p\le m$. 
\end{proof}

\begin{theorem}
If $\e_1, \e_2, \hdots, \e_n$ and $\f_1, \f_2, \hdots, f_m$ are both basis for $V$, then $m = n$. 
\end{theorem}

\begin{proof}
Both sets are linearly independent and both span $V$. Using Thm. (\ref{TH2_1}), since the $\e_i$'s are linearly independent and the $\f_i$'s span $V$,we have $n\le m$. Since the $\e_i$'s span $V$ and the $\f_i$'s are linearly 
independent, $m\le n$. Therefore $m = n$. 
\end{proof}

If $V$ possesses a basis, then all the same number of elements. We say that $V$ is {\twelveit finite dimensional} and the number of elements in the basis is called the dimension of $V$ and is denoted by $\dim V$. If 
$V$ possesses no finite basis, then we say that $V$ is {\twelveit infinite dimensional}. 

\begin{theorem}
If $V$ has dimension $n$, there cannot exist a set of more than $n$ linearly independent vectors in $V$. Any set of more than $n$ linear combinations of $n$ given vectors is always linearly dependent.\label{Th2_2}
\end{theorem}

\begin{theorem}
The maximal number of linearly independent vectors in $\R_n$ is $n$\label{Th2_3}
\end{theorem}

\begin{theorem}
If $V$ has dimension $n$, there cannot exist a set of less than $n$ vectors which span $V$. 
\end{theorem}


%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{definition}
Vectors are directed segments. The vector in $\R_n$ whose components are $a_1, a_2,...,a_n$ is denoted by $$\a = \{a_1, a_2, ..., a_n\}\hbox{~(braces)}.$$
The zero vector in $\R_n$ is denoted as $$\o = \{0,0,...,0\},$$ with $n$ components each 0. 
\end{definition}

\begin{definition}We call $p$ vectors $\a_1, \a_2, ..., \a_p$ {\bf linearly dependent} if there are $p$ numbers $\lambda_1, \lambda_2,...,\lambda_p$ not all $0$, such that 
$$\lambda_1 \a_1 + \lambda_2 \a_2 +...+ \lambda_p \a_p = \o\hbox{~ (the zero vector)}.$$
\end{definition}

If no such $p$ numbers exist, then the $p$ vectors $\a_1, \a_2, ...,\a_p$ are called {\bf linearly independent}.

\begin{theorem} 
If a subset of the $p$ vectors $\a_1, \a_2,...,\a_p$ is linearly dependent, then the set of $p$ vectors is itself linearly dependent.\label{Th1_1}
\end{theorem}

Suppose the first $r$ vectors $\a_1, \a_2, ..., \a_r, (r < p) $are linearly dependent. Then there are $r$ numbers $\lambda_1, \lambda_2, ..., \lambda_r$ not all zero such that 
$$\lambda_1 \a_1 + \lambda_2 \a_2 + ... + \lambda_r \a_r = \o.$$ Setting $\lambda_{r+1}, \lambda_{r+2}, ..., \lambda_p$ to zero we have
$$\lambda_1 \a_1 + \lambda_2 \a_2 + ... + \lambda_r \a_r + \lambda_{r+1}\a_{r+1} +... + \lambda_p \a_p = \o.$$ where the numbers $\lambda_1, \lambda_2, ..., \lambda_p$ are not all zero. Therefore the 
$p$ vectors $\a_1, \a_2, ..., \a_p$ are linearly dependent. 

\begin{theorem}
If the $p$ vectors $\a_1, \a_2, ..., \a_p$ are linearly independent, then so is every subset of these $p$ vectors. \label{Th1_2}
\end{theorem}

For otherwise, by Theorem \ref{Th1_1}, the $p$ vectors would be linearly dependent.

\begin{theorem}
If the $p$ vectors $\a_1, \a_2, ..., \a_p$ are linearly dependent and if $p>1$, then at least one of these vectors
is a linear combination of the others.\label{Th1_3}
\end{theorem}

For, there are numbers $\lambda_i$, which do not all vanish such that $$\lambda_1 \a_1 + \lambda_2 \a_2 +...+ \lambda_p \a_p = \o.$$ Suppose $\lambda_p \ne 0$. The we can solve for $\a_p$,
$$\a_p = -{\lambda_1\over \lambda_p}\a_1 -{\lambda_2\over \lambda_p}\a_2 - ... - {\lambda_{p-1}\over \lambda_p} \a_{p-1},$$ i.e. $\a_p$ is a linear combination of the vectors $\a_1, \a_2, ..., \a_{p-1}$.  

\begin{theorem}
If one of the vectors $\a_1, \a_2, ..., \a_p$ is a linear combination of the others, then vectors $\a_1,\a_2,...,\a_p$ are linearly dependent.\label{Th1_4}
\end{theorem}

Suppose $$\a_1 = \lambda_2 \a_2 + \lambda_3 \a_3 +... + \lambda_p \a_p.$$ Then we have that 
$$\a_1 - \lambda_2 \a_2 - \lambda_3 \a_3 -... - \lambda_p \a_p = \o,$$ and since the coefficient of $\a_1$ is not zero, the vectors $\a_1, \a_2, ... , \a_p$ are linearly dependent. 

\begin{theorem}
If the vectors $\a_1, \a_2, ..., \a_p$ are linearly independent, and if the vectors $\a_1, \a_2,...\a_p,\b$ are linearly dependent, then $\b$ is a linear
combination of $\a_1, \a_2, ..., \a_p$.\label{Th1_5}
\end{theorem}

We have a relation of the form $$\lambda_1 \a_1 + \lambda_2 \a_2 +... + \lambda_p \a_p + \lambda_{p+1} \b =\o,$$ where not all the $\lambda_i$ vanish. Now
$\lambda_{p+1}$ cannot be 0 because the last term would drop out and then all the other $\lambda_i$ would vanish because of the linear independence of the $\a_i$ vectors. Therefore
$\lambda_{p+1} \ne 0$ and so we have 
$$\b = -{\lambda_1\over \lambda_{p+1}}\a_1 -{\lambda_2\over \lambda_{p+1}}\a_2 - ... - {\lambda_p\over \lambda_{p+1}} \a_p.$$


\begin{theorem}
If the vectors $\a_1, \a_2, ..., \a_p$ are linearly independent, and if $\b$ is a not a linear combination of $\a_1, \a_2, ..., \a_p$, then the vectors
$\a_1, \a_2,...\a_p,\b$ are linearly independent.\label{Th1_6}
\end{theorem}

Theorem \ref{Th1_6} follows from Theorem \ref{Th1_5},\\

Let $L$ be a {\elevenit non-empty} set of vectors of $\R_n$ with the properties:
\begin{itemize}
\item[1.] If $\a$ is a vector of the set $L$, then $\lambda \a$ belongs to $L$ for every real $\lambda$.
\item[2.] If $\a$ and $\b$ are two (not necessarily distinct) vectors of the set $L$, then the vector $\a +\b$ also belongs to $L$.
\end{itemize}
Such a set of vectors, satisfying 1. and 2., is  called a {\bf vector space} in $\R_n$.

%\begin{theorem} {\bf Steinitz replacement theorem}. Let $\a_1, \a_2, ..., \a_p$ be any $p$ vectors, and let $L$ be the vector space spanned by them. Let $\b_1, \b_2, ..., \b_q$ be any $q$ linearly independent vectors of 
%$L$. Then we may replace a certain set of $q$ vectors from among the vectors $\a_1, \a_2, ..., \a_p$ by the vectors $\b_1, \b_2, ...,\b_q$ so that the remaining vectors of $\a_1, \a_2, ..., \a_p$ together with the vectors $\b_1, \b_2, ..., \b_q$ span the entire vector space $L$.\label{Th2_1}
%\end{theorem}

%\begin{proof}
%Use mathematical induction.
%\end{proof}

%\begin{theorem}
%Any set of more than $p$ linear combinations of $p$ given vectors is always linearly dependent.\label{Th2_2}
%\end{theorem}

%\begin{theorem}
%The maximal number of linearly independent vectors in $\R_n$ is $n$\label{Th2_3}
%\end{theorem}

\begin{definition}{\bf Dimension and Basis}\\
Let $L$ be any vector space in $\R_n$. The maximal number of linearly independent vectors is $\le n$. Let $p$ be this maximal number of independent vectors in $L$. We call $p$ the {\bf dimension} of $L$. Thus,
$$0 \le p \le n.$$

We call any system of $p$ linearly independent vectors of $L$, such as $\a_1, \a_2, ..., \a_p$ a {\bf basis} of $L$.
\end{definition}

\begin{theorem}
Every vector of $L$ can be represented in exactly one way as a linear combination of the basis vectors $\a_1, \a_2, ..., \a_p$.\label{Th2_4}
\end{theorem}

\begin{theorem}
The dimension of the vector space spanned by the vectors $\a_1, \a_2, ...,\a_p$ is equal to the maximal number of linearly independent vectors among $\a_1, \a_2, ..., \a_p$.\label{Th2_5}
\end{theorem}

\begin{theorem}
Any $k ~ (k\le p)$ linearly independent vectors $$\b_1, \b_2, ..., \b_k$$
 of $L$ can be extended to form a basis of $L$ by suitably adjoining to them $p-k$ other vectors $\a_{k+1}, \a_{k+2}, ..., \a_p$.\label{Th2_6}
\end{theorem}

Let $L'$ and $L''$ be any two vector spaces in $\R_n$. By the {\bf intersection} $D$ of $L'$ and $L''$ we mean the set of all those vectors belonging to both $L'$ and $L''$.\\

If we form the totality $S$ of all vectors of the form $\a' + \a''$, where $\a'$ belongs to $L'$ and $\a''$ belongs to $L''$. $S$ is called the {\bf sum} of $L'$ and $L''$.

\begin{theorem}
If the dimensions of the vector spaces $L', L'', D,$ and $S$ are respectively $p', p'', d$ and $s$, then $$p' + p'' = d + s.$$\label{Th2_7}
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\bf Subspaces }

We define a {\twelveit vector subspace } of $M$ of a  vector space $V$ to be a subset of $V$ which is closed under addition and scalar multiplication : if $\x$ and $\y$ are in $M$ then so is $\x + \y$ and so is $\lambda \x$ for any scalar $\lambda$. 

\begin{theorem}
If $V$ has a finite dimension $n$ and $M$ is a non-trivial subspace of $V$, then $M$ has finite dimension $m$ where $0 < m < n$. \label{finiteSubspace}
\end{theorem}

\begin{theorem}[Basis Reduction Theorem]
If $V = span(\v_1, \hdots, \v_m)$, then either $(\v_1, \hdots, \v_m)$ is a basis of $V$ or some $\v_i$ can be removed to obtain a basis of $V$. 
\end{theorem}
\begin{proof}
Suppose $V = span(\v_1, \hdots, \v_m)$. We start with the list $S = (\v_1, \hdots, \v_m)$, and sequentially run through all vectors $\v_k$ for $k = 1,2,\hdots,m$ to determine whether to keep or remove them from $S$:
\begin{itemize}
\item[]{\bf Step 1.} If $\v_1 = \o$, then remove $\v_1$ from $S$. Otherwise, leave $S$ unchanged.
\item[]{\bf Step k.} If $\v_k \in span(\v_1, \hdots, \v_{k-1})$, then remove $\v_k$ from $S$. Otherwise, leave $S$ unchanged. 
\end{itemize}
The final list $S$ still spans $V$ since, at each step, a vector was only discarded if it was already in the span of the previous vectors. The process also ensures that no vector is in the span of the previous vectors. 
Hence by Linear Dependence [see Definition (\ref{LinearDependence})] the final list is linearly independent. Therefore, the final list $S$ is a basis of $V$. 
\end{proof}

\begin{theorem}[Basis Extension Theorem]
If $M$ is a subspace of $V$, where $V$ is finite dimensional, then any basis of $M$ is contained in a basis of $V$. 
\end{theorem}
\begin{proof}
Suppose $V$ is finite dimensional and that $(\v_1, \hdots, \v_m)$ is linearly independent. Since $V$ is finite dimensional, there exists a finite set of vectors $(\w_1,\hdots,\w_n$ that span $V$. We want to adjoin some of the 
$\w_k$ to $(\v_1, \hdots, \v_m)$ in order to create a basis of $V$.
\begin{itemize}
\item[]{\bf Step 1.} If $\w_l \in span(\v_1, \hdots, \v_m)$, then $S = (\v_1, \hdots, \v_m$. Otherwise, $S = (\v_1, \hdots, \v_m, \w_l)$.
\item[]{\bf Step k.} If $\w_k \in span(S)$, then leave $S$ unchanged. Otherwise adjoin $\w_k$ to $S$. 
\end{itemize}
After each step, the list $S$ is still linearly independent because we only adjoined $\w_k$ if it was not in the span of the previous vectors. After $n$ steps, $\w_k\in span(S)$ for all $k = 1, 2, \hdots, n$. Since $\w_1, \hdots, \w_n)$  was a spanning set, $S$ spans $V$ so that $S$ is indeed a basis of $V$. 
\end{proof}

\begin{theorem}
If $M$ and $N$ are subspaces of a finite dimensional vector space $V$, then 
\[ \dim M + \dim N = \dim (M\cap N) + \dim(M + N).\] 
\end{theorem}

{\bf The range and null space of a linear transformation}\\

If $\theta: V \goes W$ is a linear transformation, where $V$ and $W$ are not necessarily finite dimensional, the image of $\theta$ is called the {\twelveit range} of $\theta$ and the kernel is called the {\twelveit null space}. The range and null space image and kernel of $\theta$ considered as a group homomorphism, the are subgroups of $W$ and $V$ respectively. 

\begin{theorem}
The range $R$ of $\theta: V \goes W$ is a vector subspace of $W$. 
\end{theorem}

\begin{proof}
By Thm (\ref{image}) the range is a subgroup of $W$. If $\y \in R$, so that $ \y = \theta \x$ for some $\x \in V$, $\theta(\lambda \x = \lambda \y$ and so $\lambda \y \in R$ for any scalar $\lambda$. Hence $R$ is a subspace. 
\end{proof}

\begin{theorem}
The null space $N$ of $\theta: V \goes W$ is a vector subspace of $V$. 
\end{theorem}

\begin{proof}
By Thm. (\ref{kernel}) the null space is a subgroup. If $\x \in N$, so that $\theta \x = \o$, then $\theta(\lambda\x) = \lambda \o = \o$, and hence $\lambda \x \in N$. Therefore $N$ is a subspace. 
\end{proof}

If $V$ and $W$ are finite dimensional, and \[ \dim V = n, \quad \dim W - m,\]  we see by Thm (\ref{finiteSubspace}) that the range $R$ has finite dimension $\le m$, and the null space $N$ finite dimension $\le n$. The dimension of $R$ is 
called the {\twelveit rank of } $\theta$ and that of $N$ is called the {\twelveit nullity  of } $\theta$. These are denoted by $r(\theta)$ and $\nu(\theta)$ respectively. \\

Therefore $0 \le r(\theta) \le m$ and $0 \le \nu(\theta) \le n$. If $\theta$ is the zero transformation, $r(\theta) = 0$ and $\nu(\theta) = n$, whereas if $\theta$ is an isomorphism $\r(\theta) = n$ and $\nu(\theta) = 0$. \\

In both cases rank plus nullity is $n$. and this result is generally true. 

\begin{theorem}[Rank / Nullity Theorem] If $\theta : V \goes W$ is a linear transformation between finite dimensional vector spaces with $\dim V = n$, then 
\[ r(\theta) + \nu(\theta) = n.\]
\end{theorem}